---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load library
```{r message = F, warning = F}
xlib  <- c("tidyverse", "grid", "gridExtra", "caret", "randomForest", "xgboost", "Hmisc", "rpart", "rpart.plot")
lib   <- lapply(xlib, library, character.only = TRUE) # load the required packages
```

### Load Data
```{r}
train.data  <- read_csv("data/Train_UWu5bXk.csv")
test.data   <- read_csv("data/Test_u94Q5KV.csv")
test.data$Item_Outlet_Sales <- NA

all_bms <- bind_rows(train.data, test.data)
```

### Data Cleaning
#### Missing Values: Item_Weight  
Basic R
```r
tmp <- aggregate(Item_Weight ~ Item_Identifier, data = all_bms, FUN = mean)
for (i in which(is.na(all_bms$Item_Weight))) {
  all_bms$Item_Weight[i] <- tmp$Item_Weight[tmp$Item_Identifier == all_bms$Item_Identifier[i]]
}
```

`tidyverse`
```{r all_bms1}
# Missing value of Item_Weight is filled by its average
tmp <- all_bms %>% 
  group_by(Item_Identifier) %>%
  summarise(Mean_Item_Weight = mean(Item_Weight, na.rm = T))
all_bms1 <- all_bms %>% 
  left_join(tmp) %>% 
  mutate(Item_Weight = if_else(is.na(Item_Weight), Mean_Item_Weight, Item_Weight)) %>% 
  select(-Mean_Item_Weight)

sum(is.na(all_bms1$Item_Weight))
```

#### Missing Values: Outlet_Size
##### Summary of Outlet_Size
```{r}
table(all_bms1$Outlet_Size, all_bms1$Outlet_Identifier)

all_bms1 %>%
  group_by(Outlet_Identifier, Outlet_Size) %>%
  summarise(count = n())
```

##### Fill the missing value with `tree`
```{r all_bms2}
fit <- rpart(factor(Outlet_Size) ~ Outlet_Type, 
             data = filter(all_bms, !is.na(Outlet_Size)),
             method = "class")
pred <- predict(fit, 
                data = filter(all_bms, is.na(Outlet_Size)), 
                type = "class")
all_bms2 <- all_bms1
all_bms2$Outlet_Size[is.na(all_bms2$Outlet_Size)] <- as.vector(pred)

table(all_bms2$Outlet_Size, all_bms2$Outlet_Identifier)
```

#### Classification: Unify
```{r all_bms3}
#all_bms2$Item_Fat_Content[all_bms2$Item_Fat_Content %in% 
#                           c("LF","low fat")] <- "Low Fat"
#all_bms2$Item_Fat_Content[all_bms2$Item_Fat_Content %in% 
#                           c("reg")] <- "Regular"
all_bms3 <- all_bms2 %>% 
  mutate(Item_Fat_Content = str_replace(Item_Fat_Content, "LF", "Low Fat"),
         Item_Fat_Content = str_replace(Item_Fat_Content, "low fat", "Low Fat"),
         Item_Fat_Content = str_replace(Item_Fat_Content, "reg", "Regular"))
table(all_bms3$Item_Fat_Content)
```

#### Classification: Additional
##### Information from Item_Identifier
```{r}
tmp3 <- all_bms3 %>% 
  mutate(Item_Attribute = substr(all_bms$Item_Identifier, 1, 2)) %>% 
  group_by(Item_Attribute) %>% 
  summarise(count = n())
tmp3
```

##### Food vs Non-Food  
Basic R
```r
all_bms$Item_Attribute <-  factor(substr(all_bms$Item_Identifier, 1, 2))
table(all_bms$Item_Attribute)

all_bms$Item_Fat_Content[all_bms$Item_Attribute == "NC"] <- "Non-Food"
table(all_bms$Item_Fat_Content)
```
`tidyverse`
```{r all_bms4}
all_bms4 <- all_bms3 %>% 
  mutate(Item_Attribute = substr(all_bms$Item_Identifier, 1, 2)) %>% 
  mutate(Item_Fat_Content = if_else(Item_Attribute == "NC", "Non-Food", Item_Fat_Content))
```


#### Fix Visibility == 0
There are `r sum(all_bms4$Item_Visibility==0)` itmes with `Visibility == 0`  

```{r}
tmp4 <- all_bms4 %>% 
  group_by(Outlet_Identifier) %>% 
  summarise(Item_Visibility_mean = mean(Item_Visibility))

all_bms5 <- all_bms4 %>%
  left_join(tmp4) %>% 
  mutate(Item_Visibility = if_else(Item_Visibility == 0, Item_Visibility_mean, Item_Visibility)) %>% 
  select(-Item_Visibility_mean)

sum(all_bms5$Item_Visibility==0)
```

#### New Year Column
```{r}
all_bms6 <- all_bms5 %>%
  mutate(Outlet_Years = 2013 - Outlet_Establishment_Year)
```

### Predictor: Item_Sale_Volumn
```{r}
all_bms7 <- all_bms6 %>%
  mutate(Item_Sales_Vol = round(Item_Outlet_Sales / Item_MRP + 0.5, 0))
```

### Factorize
```{r}
cols <- c("Item_Fat_Content", "Item_Type", "Outlet_Location_Type", "Outlet_Size",
          "Outlet_Type", "Outlet_Years", "Item_Attribute", "Outlet_Identifier")

all_bms8 <- all_bms7 %>% 
  mutate_at(cols, factor)
```

### Split
```{r}
all_bms <- all_bms8
train <- all_bms[!is.na(all_bms$Item_Outlet_Sales), ]
test  <- all_bms[is.na(all_bms$Item_Outlet_Sales), ]

save(train, file = "output/train_clean.RData")
save(test,  file = "output/test_clean.RData")

set.seed(1234)
ind       <- createDataPartition(train$Item_Sales_Vol, p = .7, list = FALSE)
train_val <- train[ind, ]
test_val  <- train[-ind, ]
```

### Build Model
```{r}
myformula <- Item_Sales_Vol ~ Outlet_Type + Item_Visibility + 
  Outlet_Location_Type + Item_MRP + Item_Type + Outlet_Years + Outlet_Size

model.rmse <- function(pred, act){
  sqrt(sum((act - pred)^2)/length(act))
}
```

#### Model 1: Decision Tree
```{r}
fit.tr <- rpart(myformula, 
                data = train_val, 
                method = "anova")
summary(fit.tr)
rpart.plot(fit.tr)

pred  <- predict(fit.tr, test_val)
model.rmse(pred * test_val$Item_MRP, test_val$Item_Outlet_Sales)

pred.test <- predict(fit.tr, test)
submit    <- data.frame(Item_Identifier = test.data$Item_Identifier, 
                        Outlet_Identifier = test.data$Outlet_Identifier, 
                        Item_Outlet_Sales = pred.test * test$Item_MRP)
write.csv(submit, file = "output/dtree.csv", row.names = FALSE)
```

#### Model 2: Random Forest
```{r}
set.seed(2345)
fit.rf  <- randomForest(myformula, 
                       data = train_val,
                       ntree = 500)
summary(fit.rf)
pred    <- predict(fit.rf, test_val)
model.rmse(pred*test_val$Item_MRP, test_val$Item_Outlet_Sales)

pred.test <- predict(fit.rf, test)
submit    <- data.frame(Item_Identifier = test.data$Item_Identifier, 
                        Outlet_Identifier = test.data$Outlet_Identifier, 
                        Item_Outlet_Sales = pred.test * test$Item_MRP)
write.csv(submit, file = "output/rf.csv", row.names = FALSE)
```

#### Model 3: GBM
```{r}
Ctrl    <- trainControl(method = "repeatedcv", number = 5, repeats = 5)
set.seed(3456)
fit.gbm <- train(myformula, 
                 data = train_val, 
                 trControl = Ctrl, 
                 method = "gbm", 
                 verbose = FALSE)
summary(fit.gbm)
pred    <- predict(fit.gbm, test_val)
model.rmse(pred*test_val$Item_MRP, test_val$Item_Outlet_Sales)

pred.test <- predict(fit.gbm, test)
submit    <- data.frame(Item_Identifier = test$Item_Identifier, 
                        Outlet_Identifier = test$Outlet_Identifier, 
                        Item_Outlet_Sales = pred.test*test$Item_MRP)
write.csv(submit, file = "output/gbm5cv.csv", row.names = FALSE)
```

#### Model 4: xgboost
##### Step 1: Sparse Matrix
```{r}
mymatrix    <- function(train){
matrix_num  <- train[ , c("Item_Visibility", "Item_MRP")]
matrix_num  <- cbind(matrix_num,
                      model.matrix(~ Outlet_Type - 1, train),
                      model.matrix(~ Outlet_Location_Type - 1, train),
                      model.matrix(~ Outlet_Size - 1, train),
                      model.matrix(~ Item_Type - 1, train),
                      model.matrix(~ Outlet_Years - 1, train)
                      )
  return(data.matrix(matrix_num))
}


xgb.train_val <- mymatrix(train_val)
xgb.test_val  <- mymatrix(test_val)
xgb.test      <- mymatrix(test)

dtrain_val  <- xgb.DMatrix(data = xgb.train_val, label=train_val$Item_Sales_Vol)
dtest_val   <- xgb.DMatrix(data = xgb.test_val, label=test_val$Item_Sales_Vol)
dtest_sub   <- xgb.DMatrix(data = xgb.test)
```

##### Prelimilary
```{r}
model     <- xgboost(data = dtrain_val, nround = 5)

summary(model)
pred      <- predict(model,dtest_val)
model.rmse(pred*test_val$Item_MRP, test_val$Item_Outlet_Sales)
xgb.importance(colnames(xgb.train_val), model)

pred.test <- predict(model, dtest_sub)
submit    <- data.frame(Item_Identifier = test$Item_Identifier, 
                        Outlet_Identifier = test$Outlet_Identifier,
                        Item_Outlet_Sales = pred.test * test$Item_MRP)
write.csv(submit, file = "output/xgb.csv", row.names = FALSE)
```

##### Optimization
```{r}
model_tuned <- xgboost(data = dtrain_val,
                       nround = 10,
                       max.depth = 5
                       )
summary(model_tuned)
pred <- predict(model_tuned,dtest_val)
model.rmse(pred*test_val$Item_MRP, test_val$Item_Outlet_Sales)
xgb.importance(colnames(xgb.train_val), model_tuned)

pred.test <- predict(model_tuned, dtest_sub)
submit <- data.frame(Item_Identifier = test$Item_Identifier, 
                     Outlet_Identifier = test$Outlet_Identifier, 
                     Item_Outlet_Sales = pred.test * test$Item_MRP)
write.csv(submit, file = "output/xgbn10d5.csv", row.names = FALSE)
```


